Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-03-13 paper/paper.tex
revisions made

The underlying abstract data type and the universal algorithmic pattern found in STV characterise components of a finite state machine which we call \emph{generic STV}~(Definition\ref{STVMachine}). The set of machine states consist of the discrete states of computation $\mathcal{S}$ mentioned earlier. The seven generic actions form the transition labels $\mathcal{T}$ of the machine. Last, the everywhere present pre- and post-conditions which define (content of)  the actions, make a small-step operational semantics for the machine.  
Note that the states of computation recorded in all possible  evidence now become identical with machine states.

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ where  for $i\in\{1...j_{t}\}$, $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV)~\cite{vic}, distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such precondition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in postconditions. For example transfer-elected of the ACT STV~\cite{act} used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

Also there are some \emph{implicit} conditions present in legal documents describing transfer-elect. These constraints come to attention either as the result of a straightforward understanding of the explicit conditions above or as   auxiliary components that are taken for granted by legislators but are
necessary for proper functioning of the explicit constraints;
\begin{enumerate}
\item every candidate in the pre- and post-state of transfer-elected has a unique tally and pile
\item no one is elected or eliminated by applying transfer-elect
\item no candidate attracts any new vote by transfer-elect and therefore tallies remain the same
\item candidates whose names appears in the backlog of elected are indeed among elected candidates
\item any elected candidate is no longer a continuing candidate so that they do not receive votes any further
\item the list of competing candidates in the election is not empty and has no duplication of names
\end{enumerate}  
Conjunctions of formal declarations of the above explicit and implicit conditions forms the semantics of the transition label transfer-elect.   The predicate TransferAuxSpec defines  the semantics of this transition.
 \begin{small}
\begin{holthmenv}
  \HOLConst{TransferAuxSpec}\;(\HOLFreeVar{qu}\HOLSymConst{,}\HOLFreeVar{st}\HOLSymConst{,}\HOLFreeVar{l})\;\HOLFreeVar{ba}\;\HOLFreeVar{t}\;\HOLFreeVar{t\sp{\prime}}\;\HOLFreeVar{p}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{bl}\;\HOLFreeVar{bl\sb{\mathrm{2}}}\;\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\;\HOLFreeVar{e}\;\HOLFreeVar{e\sp{\prime}}\;\HOLFreeVar{h}\;\HOLFreeVar{h\sp{\prime}}\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLFreeVar{bl\sb{\mathrm{2}}}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{bl}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{ba}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{t\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{t}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{e\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{e}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLFreeVar{h\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{length}\;\HOLFreeVar{e}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{st}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{d}.\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{d}\;(\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenDoublePlus}\;\HOLFreeVar{e})\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{d}\;\HOLFreeVar{l})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{d}.\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{d}\;\HOLFreeVar{bl}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{d}\;\HOLFreeVar{l})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;(\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenDoublePlus}\;\HOLFreeVar{e})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{Valid_PileTally}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally}\;\HOLFreeVar{p}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\\;\;\HOLConst{ValidInitCandList}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}
\;\;\HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{p})
\end{holthmenv}
\end{small}

Transfer-elect, as is the case for other formalised transitions, is parametrised by the quota $qu$, number of initial vacancies $st$ and the list $l$ of all candidates competing in the election. The semantics of the transition declares that given the ballots $ba$, tally $t$, pile $p$, backlog of elected $bl$, backlog of eliminated $bl_{2}$ and the list of elected $e$ and continuing candidates $h$ in the pre-state of transfer-elected, and their respective counterparts in the post-state characterised by having a prime symbol, some conditions as specified above are satisfied by the transition.

@@

\;\;\HOLKeyword{if}\;\HOLConst{\HOLConst{mem}}\;\HOLFreeVar{l\sb{\mathrm{0}}}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLKeyword{then}\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLFreeVar{ls}\\

@@

Conjunctions of the computational implementations  define a computational twin TransferAuxDec for the specification  of the transfer-elect semantics TransferAuxSpec. 

@@

\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{ba}\;\HOLBoundVar{nba}\;\HOLBoundVar{t}\;\HOLBoundVar{nt}\;\HOLBoundVar{p}\;\HOLBoundVar{np}\;\HOLBoundVar{bl}\;\HOLBoundVar{nbl}\;\HOLBoundVar{bl\sb{\mathrm{2}}}\;\HOLBoundVar{nbl\sb{\mathrm{2}}}\;\HOLBoundVar{e}\;\HOLBoundVar{ne}\;\HOLBoundVar{h}\;\HOLBoundVar{nh}.\\
\;\;\;\;\;\;\HOLFreeVar{j\sb{\mathrm{1}}}\;\HOLSymConst{=}\;\HOLConst{NonFinal}\;(\HOLBoundVar{ba}\HOLSymConst{,}\HOLBoundVar{t}\HOLSymConst{,}\HOLBoundVar{p}\HOLSymConst{,}\HOLBoundVar{bl}\HOLSymConst{,}\HOLBoundVar{bl\sb{\mathrm{2}}}\HOLSymConst{,}\HOLBoundVar{e}\HOLSymConst{,}\HOLBoundVar{h})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\HOLConst{TransferAuxSpec}\;(\HOLFreeVar{qu}\HOLSymConst{,}\HOLFreeVar{st}\HOLSymConst{,}\HOLFreeVar{l})\;\HOLBoundVar{ba}\;\HOLBoundVar{t}\;\HOLBoundVar{nt}\;\HOLBoundVar{p}\;\HOLBoundVar{np}\;\HOLBoundVar{bl}\;\HOLBoundVar{bl\sb{\mathrm{2}}}\;\HOLBoundVar{nbl\sb{\mathrm{2}}}\;\HOLBoundVar{e}\;\HOLBoundVar{ne}\;\HOLBoundVar{h}\\
\;\;\;\;\;\;\;\;\HOLBoundVar{nh}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{c\sp{\prime}}.\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c\sp{\prime}}\;\HOLBoundVar{h}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{x}.\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{c\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{x})\;\HOLBoundVar{t}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{x}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{qu})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l\sp{\prime}}\;\HOLBoundVar{c}.\\
\;\;\;\;\;\;\;\;\;\;(\HOLBoundVar{bl}\;\HOLSymConst{=}\;\HOLBoundVar{c}\HOLSymConst{::}\HOLBoundVar{l\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{l\sp{\prime\prime}}.\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{c}\HOLSymConst{,}\HOLBoundVar{l\sp{\prime\prime}})\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{l\sp{\prime\prime}}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLConst{\ensuremath{[\,]}})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{nbl}\;\HOLSymConst{=}\;\HOLBoundVar{l\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\;\;\;\;\HOLBoundVar{nba}\;\HOLSymConst{=}\;\HOLConst{last}\;(\HOLConst{get_cand_pile}\;\HOLBoundVar{c}\;\HOLBoundVar{p})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{c}\HOLSymConst{,}\HOLConst{\ensuremath{[\,]}})\;\HOLBoundVar{np}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{d\sp{\prime}}.\\
\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLBoundVar{d\sp{\prime}}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{l\sp{\prime}}.\\
\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;(\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{l\sp{\prime}})\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{l\sp{\prime}})\;\HOLBoundVar{np})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;(\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{l\sp{\prime}})\;\HOLBoundVar{np}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{l\sp{\prime}})\;\HOLBoundVar{p}))\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\;\;\;\HOLFreeVar{j\sb{\mathrm{2}}}\;\HOLSymConst{=}\;\HOLConst{NonFinal}\;(\HOLBoundVar{nba}\HOLSymConst{,}\HOLBoundVar{nt}\HOLSymConst{,}\HOLBoundVar{np}\HOLSymConst{,}\HOLBoundVar{nbl}\HOLSymConst{,}\HOLBoundVar{nbl\sb{\mathrm{2}}}\HOLSymConst{,}\HOLBoundVar{ne}\HOLSymConst{,}\HOLBoundVar{nh})

@@

We augment the formal counterparts of the $\bullet_{i}$ conditions  to the clauses given in section~\ref{sec:MachineSem} to obtain the specification TransferActSpec for transfer-elect of ACT STV.  The last conjunct of TransferActSpec asserts that the pile of every candidate other than $c$ remains the same in both pre- and post-state of the transition. Also note that we place the last parcel of the pile of the candidate $c$ whose votes are transferred first into the list of uncounted ballots so that in the subsequent transition count deals with distributing the votes and ``recalculating'' candidates' tallies and piles. Finally, because of efficiency purposes our system is designed to update the fractional transfer value of the surplus in formalisation of the elect transition instead of the transfer-elected transition.

@@

For example, the function get\textunderscore{}cand\textunderscore{}tally looks through a tally list $t$ and finds the tally of an input candidate name $c$. We then verify this function  computes the tally of candidates correctly and that indeed every candidate is assigned only one tally (item 1 of the implicit machine conditions).

@@

  \HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLFreeVar{c}\HOLSymConst{,}\HOLFreeVar{x})\;\HOLFreeVar{t}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{get_cand_tally}\;\HOLFreeVar{c}\;\HOLFreeVar{t}\;\HOLSymConst{=}\;\HOLFreeVar{x}

@@

Using this function, we implement another function less\textunderscore{}than\textunderscore{}quota which checks if the tally of every candidate in a given list $ls$ is below the quota (item $\bullet_{1}$).

@@

We show less\textunderscore{}than\textunderscore{}quota is a  computational realisation of  its specification: 

@@

Moreover less\textunderscore{}than\textunderscore{}quota enforces its specification.

@@

The clause $\star_{4}$ itself consists of the following sub-clauses.

>>>>

The underlying abstract data type and the universal algorithmic pattern found in STV characterise components of a finite state machine which we call \emph{generic STV}~(Definition\ref{STVMachine}). The set of machine states consist of the discrete states of computation $\mathcal{S}$ mentioned earlier. The seven generic actions given in Definition~\ref{transitions} form the transition labels $\mathcal{T}$ of the machine. Last, the everywhere present pre- and post-conditions which define (content of)  the actions, make a small-step operational semantics for the machine.  
Note that the states of computation recorded in all possible  evidence now become identical with machine states.

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ 
 where $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV)~\cite{vic}, distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such precondition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in postconditions. For example transfer-elected of the ACT STV~\cite{act} used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

Also there are some \emph{implicit} conditions present in legal documents describing transfer-elect. These constraints come to attention either as the result of a straightforward understanding of the explicit conditions above or as   auxiliary components that are taken for granted by legislators but are
necessary for proper functioning of the explicit constraints;
\begin{enumerate}
\item every candidate in the pre- and post-state of transfer-elected has a unique tally and pile
\item no one is elected or eliminated by applying transfer-elect
\item no candidate attracts any new vote by transfer-elect and therefore tallies remain the same
\item candidates whose names appears in the backlog of elected are indeed among elected candidates
\item any elected candidate is no longer a continuing candidate so that they do not receive votes any further
\item the list of competing candidates in the election is not empty and has no duplication of names
\end{enumerate}  
Conjunctions of formal declarations of the above explicit and implicit conditions form the semantics of the transition label transfer-elect.   The predicate TransferAuxSpec defines  the semantics of this transition.
 \begin{small}
\begin{holthmenv}
  \HOLConst{TransferAuxSpec}\;(\HOLFreeVar{qu}\HOLSymConst{,}\HOLFreeVar{st}\HOLSymConst{,}\HOLFreeVar{l})\;\HOLFreeVar{ba}\;\HOLFreeVar{t}\;\HOLFreeVar{t\sp{\prime}}\;\HOLFreeVar{p}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{bl}\;\HOLFreeVar{bl\sb{\mathrm{2}}}\;\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\;\HOLFreeVar{e}\;\HOLFreeVar{e\sp{\prime}}\;\HOLFreeVar{h}\;\HOLFreeVar{h\sp{\prime}}\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLFreeVar{bl\sb{\mathrm{2}}}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{bl}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{ba}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{t\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{t}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{e\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{e}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLFreeVar{h\sp{\prime}}\;\HOLSymConst{=}\;\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{length}\;\HOLFreeVar{e}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{st}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{d}.\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{d}\;(\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenDoublePlus}\;\HOLFreeVar{e})\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{d}\;\HOLFreeVar{l})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{d}.\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{d}\;\HOLFreeVar{bl}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{d}\;\HOLFreeVar{l})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;(\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenDoublePlus}\;\HOLFreeVar{e})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{Valid_PileTally}\;\HOLFreeVar{t}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally}\;\HOLFreeVar{p}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{Valid_PileTally}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\\;\;\HOLConst{ValidInitCandList}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}
\;\;\HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{p})
\end{holthmenv}
\end{small}

Transfer-elect, as is the case for other formalised transitions, is parametrised by the quota $qu$, number of initial vacancies $st$ and the list $l$ of all candidates competing in the election. The semantics of the transition declares that given the ballots $ba$, tally $t$, pile $p$, backlog of elected $bl$, backlog of eliminated $bl_{2}$ and the list of elected $e$ and continuing candidates $h$ in the pre-state of transfer-elected, and their respective counterparts in the post-state characterised by having a prime symbol, some conditions as specified above are satisfied by the transition.  Note that \textsf{mem} in Definition of TransferAuxSpec declares list membership.

@@

\;\;\HOLKeyword{if}\;\HOLConst{\HOLConst{mem}}\;\HOLFreeVar{l\sb{\mathrm{0}}}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLKeyword{then}\;\HOLConst{Valid_PileTally_dec2}\;\HOLFreeVar{t}\;\HOLFreeVar{ls}\;

@@

Conjunctions of the computational implementations  define a computational twin TransferAuxDec for the specification  of the transfer-elected semantics TransferAuxSpec. 

@@

\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{ba}\;\HOLBoundVar{ba\sp{\prime}}\;\HOLBoundVar{t}\;\HOLBoundVar{t\sp{\prime}}\;\HOLBoundVar{p}\;\HOLBoundVar{p\sp{\prime}}\;\HOLBoundVar{bl}\;\HOLBoundVar{bl\sp{\prime}}\;\HOLBoundVar{bl\sb{\mathrm{2}}}\;\HOLBoundVar{bl\sb{\mathrm{2}}\sp{\prime}}\;\HOLBoundVar{e}\;\HOLBoundVar{e\sp{\prime}}\;\HOLBoundVar{h}\;\HOLBoundVar{h\sp{\prime}}.\\
\;\;\;\;\HOLFreeVar{j\sb{\mathrm{1}}}\;\HOLSymConst{=}\;\HOLConst{NonFinal}\;(\HOLBoundVar{ba}\HOLSymConst{,}\HOLBoundVar{t}\HOLSymConst{,}\HOLBoundVar{p}\HOLSymConst{,}\HOLBoundVar{bl}\HOLSymConst{,}\HOLBoundVar{bl\sb{\mathrm{2}}}\HOLSymConst{,}\HOLBoundVar{e}\HOLSymConst{,}\HOLBoundVar{h})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\HOLConst{TransferAuxSpec}\;(\HOLFreeVar{qu}\HOLSymConst{,}\HOLFreeVar{st}\HOLSymConst{,}\HOLFreeVar{l})\;\HOLBoundVar{ba}\;\HOLBoundVar{t}\;\HOLBoundVar{t\sp{\prime}}\;\HOLBoundVar{p}\;\HOLBoundVar{p\sp{\prime}}\;\HOLBoundVar{bl}\;\HOLBoundVar{bl\sb{\mathrm{2}}}\;\HOLBoundVar{bl\sb{\mathrm{2}}\sp{\prime}}\;\HOLBoundVar{e}\;\HOLBoundVar{e\sp{\prime}}\;\HOLBoundVar{h}\HOLBoundVar{h\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\\\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{c\sp{\prime}}.\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c\sp{\prime}}\;\HOLBoundVar{h}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{x}.\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{c\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{x})\;\HOLBoundVar{t}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{x}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{qu})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l\sb{\mathrm{1}}}\;\HOLBoundVar{c}.
\;(\HOLBoundVar{bl}\;\HOLSymConst{=}\;\HOLBoundVar{c}\HOLSymConst{::}\HOLBoundVar{l\sb{\mathrm{1}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{l\sb{\mathrm{2}}}.\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{c}\HOLSymConst{,}\HOLBoundVar{l\sb{\mathrm{2}}})\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{l\sb{\mathrm{2}}}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLConst{\ensuremath{[\,]}})\\\;\;\;\;\;\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{bl\sp{\prime}}\;\HOLSymConst{=}\;\HOLBoundVar{l\sb{\mathrm{1}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{ba\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{last}\;(\HOLConst{getCandPile}\;\HOLBoundVar{c}\;\HOLBoundVar{p})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{c}\HOLSymConst{,}\HOLConst{\ensuremath{[\,]}})\;\HOLBoundVar{p\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\HOLSymConst{*}\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{d}.
\;\HOLBoundVar{d}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{l\sb{\mathrm{3}}}.\;(\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d}\HOLSymConst{,}\HOLBoundVar{l\sb{\mathrm{3}}})\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d}\HOLSymConst{,}\HOLBoundVar{l\sb{\mathrm{3}}})\;\HOLBoundVar{p\sp{\prime}})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\HOLSymConst{*}\;\;\;(\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d}\HOLSymConst{,}\HOLBoundVar{l\sb{\mathrm{3}}})\;\HOLBoundVar{p\sp{\prime}}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d}\HOLSymConst{,}\HOLBoundVar{l\sb{\mathrm{3}}})\;\HOLBoundVar{p}))\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\;\;\;\HOLFreeVar{j\sb{\mathrm{2}}}\;\HOLSymConst{=}\;\HOLConst{NonFinal}\;(\HOLBoundVar{nba}\HOLSymConst{,}\HOLBoundVar{t\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{p\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{bl\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{bl\sb{\mathrm{2}}\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{e\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{h\sp{\prime}})

@@

We augment the formal counterparts of the $\bullet_{i}$ conditions  to the clauses given in Section~\ref{sec:MachineSem} to obtain the specification TransferActSpec for transfer-elect of ACT STV.  
%The last conjunct of TransferActSpec asserts that the pile of every candidate other than $c$ remains the same in both pre- and post-state of the transition. 
The lines denoted by the star symbol~($\star$) in Definition of TransferActSpec formalise the Clause~$\bullet_{7}$ above. Also note that we place the last parcel of the pile of the candidate $c$ whose votes are transferred first into the list of uncounted ballots so that in the subsequent transition count deals with distributing the votes and ``recalculating'' candidates' tallies and piles. Finally, because of efficiency purposes our system is designed to update the fractional transfer value of the surplus in formalisation of the elect transition instead of the transfer-elected transition.

@@

For example, the function getCandTally looks through a tally list $t$ and finds the tally of an input candidate name $c$. We then verify this function  computes the tally of candidates correctly and that indeed every candidate is assigned only one tally (item 1 of the implicit machine conditions).

@@

  \HOLConst{distinct}\;(\HOLConst{map}\;\HOLConst{fst}\;\HOLFreeVar{t})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLFreeVar{c}\HOLSymConst{,}\HOLFreeVar{x})\;\HOLFreeVar{t}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{getCandTally}\;\HOLFreeVar{c}\;\HOLFreeVar{t}\;\HOLSymConst{=}\;\HOLFreeVar{x}

@@

Using this function, we implement another function lessThanQuota which checks if the tally of every candidate in a given list $ls$ is below the quota (item $\bullet_{1}$).

@@

We show that the function lessThanQuota is a  computational realisation of  its specification: 

@@

Moreover lessThanQuota enforces its specification.

@@

The Clause $\star_{4}$ itself consists of the following sub-clauses.
